#Loading the Required Libraries
```{r}
library(tidyverse)
library(skimr)
library(DataExplorer)
library(stringr)
library(patchwork)
library(ggrepel)
```


#Loading the Data
```{r}
cars <- read_csv("vehicle_data/car_prices.csv")

```



```{r}
head(cars)

```


```{r}
dim(cars)
```

```{r}
introduce(cars)
```



```{r}
plot_str(cars)

```

```{r}
theme <- theme(plot.margin = margin(10, 10, 10, 5, "mm"),              
              plot.background = element_rect(fill = "#f6f5f4"), 
              plot.title = element_text(size = 24),                   
              plot.subtitle = element_text(size = 16),                
              axis.text = element_text(size = 16),
              axis.title = element_text(size = 16),              
              legend.title = element_text(size = 16),                 
              legend.text = element_text(size = 14))
```

```{r}
nrow(cars)
head(cars, 3)
```



```{r}
skim(cars)
```


```{r}
plot_intro(cars)
```

```{r}
create_report(cars)
```




```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
plot_missing(
  cars,
  group = list(Good = 0.002, OK = 0.05, Bad = 0.8, Remove = 1),
  missing_only = FALSE, # true if you want to show only the columns with missing values (important if you have high number of features with no missing values)
  geom_label_args = list(),
  title = 'Missing values in each column',
  ggtheme = theme,
  theme_config = list(legend.position = c("top"))
)
```

```{r}
unique_values <- rapply(cars,function(x) length(unique(x)))
unique_values

```


```{r}
plot_bar(cars)
```





```{r}

# duplicated(cars$vin)
length(cars$vin[duplicated(cars$vin)])
cars$vin[duplicated(cars$vin)]
```

```{r}
filter(cars, vin == ('2cndl13f056137366') | vin == ('waulfafh7dn005779') |vin == ('1ftsw21r48ed06505'))
```


```{r}
unique(cars$year)
```

```{r}
unique(cars$make)
```

```{r}
unique(cars$model)

```

```{r}
unique(cars$transmission)

```

```{r}
unique(cars$color)

```

```{r}
unique(cars$body)

```

```{r}
unique(cars$condition)

```

```{r}
unique(cars$interior)
```

```{r}
filter(cars,interior=="—")

```



```{r}
unique(cars$color)
```

filtering the "—" in color and interior
```{r}

cars <- cars %>% mutate(interior=gsub("—","Unknown",cars$interior)) %>%  mutate(color=gsub("—","Unknown",cars$color))

```


```{r}
unique(cars$interior)
unique(cars$interior)
```


```{r}
unique(cars$state)
```
filtering state
```{r}
faulty_states<-cars %>% filter(str_detect(state,"3vw"))
nrow(faulty_states)
```


```{r}
cars <- cars[!grepl("3vw",cars$state),]
```



```{r}
unique(cars$state)
```


```{r}
cars$state=toupper(cars$state)
unique(cars$state)
```


```{r}
length(grepl("sfvver",cars$state))
```


```{r}
cars.years <- cars
dim(cars)
```


```{r}

cars<-na.omit(cars)
dim(cars)

```


```{r}
plot_missing(cars)
```



```{r}
cars$make <- str_to_title(cars$make)
cars$model <- str_to_title(cars$model)
cars$body <- str_to_title(cars$body)
```


```{r}
head(cars)
```

#EDA and Visualization

```{r}
mean(cars$sellingprice)
max(cars$sellingprice)
min(cars$sellingprice)
```


### checking the car details with max selling price
```{r}
filter(cars,sellingprice==230000)
```


checking the car details with min price($1)
```{r}
filter(cars,sellingprice<=1)

```



```{r}
# options(repr.plot.width =5, repr.plot.height = 6)
# hist <- ggplot(cars, aes(x = sellingprice)) + 
#         geom_histogram(binwidth=2000, fill='#FF9999')+
#         labs(title ="Sale Price distribution. ")
# box <- ggplot(cars, aes( y = sellingprice ) ) +
#         geom_boxplot(fill='#4DE5B1') + 
#         labs(title ="Boxplot of Sale Price")
# hist+box

```


```{r}
plot_histogram(cars$sellingprice,ggtheme = theme,title="Sale Price Distribution",geom_histogram_args =list(bins=100))
ggplot(cars, aes( y = sellingprice ) ) +
        geom_boxplot(fill='#4DE5B1') + 
        labs(title ="Boxplot of Sale Price") +
        theme

```



```{r}
# options(repr.plot.width =50, repr.plot.height = 12)

bar <- ggplot(cars, aes(x = transmission, fill = transmission )) + 
        geom_bar(stat ='count', color='#000000' , alpha=0.99)+
        geom_text(aes(label=..count..),stat='count', position=position_dodge(0.9),vjust=-0.5, color='black')+
        labs(title="Transmission System", subtitle ="Automatic cars are the majority. Afterall, we are dealing with\nAmerican car owners. No surprise here." )+
        theme +
        theme(axis.text.x = element_text(angle = 0, hjust=0.5))

box1 <- ggplot(cars, aes(x = transmission, y = sellingprice, fill = transmission)) +
        geom_boxplot()+
        labs(title="Sale price: full scale", subtitle ="Relatively more 'outliers' in\nthe Automatic cars category" )+
        theme

box2 <- ggplot(cars, aes(x = transmission, y = sellingprice, fill = transmission)) +
        geom_boxplot()+
        labs(title="Sale price: Zoomed in", subtitle ="Automatic cars sold for slightly\n higher price." )+
        theme + ylim(0, 30000)
par(mfrow=c(1,1))             
bar
par(mfrow=c(1,2))
box1
box2
```




### Sale Price on Vehicle MMR
```{r}

options(repr.plot.width = 12, repr.plot.height = 6)
ggplot(cars, aes(x =mmr, y =sellingprice, na.rm =TRUE)) + 
        geom_point(color='#f00000')+
        labs(title='MMR vs SalePrice', subtitle = 'The MMR values are well in agreedment with the SalePrice\nexcept few outliers!')+
        geom_smooth(method=lm,color="black")+
        theme
```





```{r}
filter(cars, sellingprice > 100000 & mmr < 25000)
filter(cars, (sellingprice < 50000) & (mmr > 100000))

```



```{r}

options(repr.plot.width = 20, repr.plot.height = 8)
df <- filter(cars, sellingprice > 150000)

p10<-ggplot(df, aes(x = year, y =sellingprice, label = make)) + 
        geom_point(color='#f00000')+
        geom_text_repel(aes(label = make), size=5.5)+
        labs(title='High-priced selected cars', subtitle='That Ford car looks out of place indeed. The rest are luxury brands.')+
        annotate(geom = "curve", x = 2013, y = 220000, xend = 2014, yend = 230000, curvature = -0.25,   
                 arrow = arrow(length = unit(3, "mm")))+
        annotate(geom='text', x=2013, y=215000, label='A weird outlier indeed!', size=6, color='navyblue')+       
        theme 
p10

```


### Sale Price on Odometer 
```{r}
options(repr.plot.width = 12, repr.plot.height = 6)
ggplot(cars, aes(x =odometer, y =sellingprice, na.rm =TRUE)) + 
        geom_point(color='#f00000')+
        geom_smooth(formula = y ~ log(x))+
        labs(title='Odometer vs Sale Price')+
        theme
```



### Sale Price on Vehicle Condition

```{r}
options(repr.plot.width = 12, repr.plot.height = 6)
ggplot(cars, aes(x =condition, y =sellingprice)) + 
        geom_point(color='#F8766D')+
        geom_smooth(formula = y ~ x, method = "lm")+
        labs(title= 'Cars condition on Sale Price', 
            subtitle = 'There is a trend in car condition and sale price.' )+
        theme
```





### Plotting selling price vs condition on slected cars and selected price range

```{r}
cars_ford_ <- filter(cars, make=='Ford' & model=='Fusion' & year=='2012' & body=='Sedan' & odometer <= 25000)
cars_toyota_ <- filter(cars, make=='Toyota' & model=='Camry' & year=='2012' & body=='Sedan' & odometer <= 25000)

options(repr.plot.width = 16, repr.plot.height = 6)
f <- ggplot(cars_ford_, aes(x =condition, y =sellingprice)) + 
        geom_point(color='seagreen')+
        geom_smooth(formula = y ~ x, method = "lm", color='red')+
        labs(title= 'Cars condition on Sale Price', 
            subtitle ="Ford |Fusion |Sedan |2012 |<= 25K km")+
        theme
t <- ggplot(cars_toyota_, aes(x =condition, y =sellingprice)) + 
        geom_point(color='gold')+
        geom_smooth(formula = y ~ x, method = "lm", color='red')+
        labs(title= 'Cars condition on Sale Price', 
            subtitle ="Toyota |Camry |Sedan |2012 |<= 25K km")+
        theme
f
t
```




### Average Sale price  based on Year

```{r}
options(repr.plot.width = 12, repr.plot.height = 6)

df <- cars.years %>% group_by(year) %>%
        summarise(Avg.SalePrice = mean(sellingprice))

ggplot(df, aes(x = year, y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', alpha=0.6, color='gold')+
        labs(title="Year on Sale price" )+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))

```






### frequency plot based on Year
```{r}
cars.years %>% 
    group_by(year) %>% 
    filter(!is.na(year)) %>%
    summarise(count = n()) %>% 
    ggplot(aes(x = reorder(year,(-count)), y = count)) + 
        geom_bar(stat = 'identity', fill='salmon', color='black') + 
        labs(subtitle="Given the dataset is collected in 2015, the most recent cars\nare likely to be more common as well as high sale on value")+
        xlab('Year')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))
```






```{r}
filter(cars.years, year < 1983)

```



```{r}
unique(cars$make)
```




### Renaming the make names

```{r}
cars_new <- cars %>%
    mutate(make = replace(make, (make == 'Mercedes-B'| make == 'Mercedes'), 'Mercedes-Benz')) %>%
    mutate(make = replace(make, make == "Vw" ,"Volkswagen"))%>%
    mutate(make = replace(make, make == "Landrover", "Land Rover"))%>%
    mutate(make = replace(make, (make == "Ford Truck" | make == "Ford Tk"), "Ford"))%>%
    mutate(make = replace(make, make == "Chev Truck", "Chevrolet"))%>%
    mutate(make = replace(make, make == "Mazda Tk", "Mazda"))%>%
    mutate(make = replace(make, make == "Gmc Truck", "Gmc"))%>%
    mutate(make = replace(make, (make == "Dodge Tk"), "Dodge"))%>%
    mutate(make = replace(make, make == "Hyundai Tk", "Haundai"))
```




### Plotting  the most expensive Brands and Popular models

```{r}
options(repr.plot.width = 30, repr.plot.height = 10)

df <- cars_new %>% group_by(make) %>%
      summarise(Avg.SalePrice = mean(sellingprice)) %>%
      arrange(desc(Avg.SalePrice)) %>%
      top_n(15)


p1<-ggplot(df, aes(x = Avg.SalePrice,y = reorder(make, Avg.SalePrice))) +
        geom_segment(aes(x = 0, reorder(make, Avg.SalePrice), xend = Avg.SalePrice, yend = make), color = "blue", thickness=10) +
        geom_point( size=10, color="blue", fill=alpha("orange", 0.3), alpha=0.7, shape=20, stroke=3)+
        labs(title="Top 15 expensive car brands (avg.)", subtitle ='No surprise as the top three most expensive cars are Ferrari, Bentley, and\nPorsche. Lexury brands are expensive even in thier second life.')+
        xlab('Sale Price') + ylab('Make')+
        geom_point() + theme 

df <- cars_new %>% group_by(make) %>% 
      summarise(freq = n()) %>% 
      arrange(desc(freq)) %>%
      top_n(15) 

p2<-ggplot(df, aes(y = reorder(make,(freq)), x = freq)) +
        geom_segment(aes(x = 0, reorder(make, freq), xend = freq, yend = make), color = "salmon", thickness=10) +
        geom_point( size=10, color="salmon", fill=alpha("orange", 0.3), alpha=0.7, shape=20, stroke=3)+
        labs(title="Top 15 popular car brands ", subtitle ='No surprise as the top three most expensive cars are Ferrari, Bentley, and\nPorsche. Lexury brands are expensive even in thier second life.')+
        xlab('count') + ylab('Make')+
        geom_point() + theme 
p1
p2
```






```{r}
cars_ford <- filter(cars, make=='Ford' )
cars_chev <- filter(cars, make=='Chevrolet')
cars_toyota <- filter(cars, make=='Toyota' )
cars_luxury <- filter(cars, (make=='Rolls-Royce' | make=='Ferrari'| make=='Lamborghini'))
```





```{r}
options(repr.plot.width = 18, repr.plot.height = 10)

df <- cars_luxury %>% group_by(model) %>%
      summarise(Avg.SalePrice = mean(sellingprice))%>%
      arrange(desc(Avg.SalePrice))%>%
      top_n(15) 

p1<-ggplot(df, aes(y = reorder(model, Avg.SalePrice), x = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'black', fill= 'blue')+
        labs(title="Lexury Cars Model vs Sale price" )+ ylab('Model') +
        theme +
        theme(axis.text.x = element_text(angle = 0, hjust=0.5))

df_ford <- cars_luxury %>% group_by(model) %>% 
           summarise(freq = n()) %>% 
           arrange(desc(freq))%>%
           top_n(15)
    
p2<-ggplot(df_ford, aes(y = reorder(model,(freq)), x = freq)) + 
        geom_bar(stat = 'identity', fill="seagreen", color='black') + 
        labs(title="Luxury Cars Model", subtitle= "")+ ylab('Model')+
        xlab('count')+
        theme + theme(axis.text.x = element_text(angle = 0, hjust=0.5))

###

df_pop <- cars_luxury %>% group_by(body) %>% 
            summarise(freq = n())  
    
p3<-ggplot(df_pop, aes(y = reorder(body,(freq)), x = freq)) + 
        geom_bar(stat = 'identity', color="#000666", fill='gold') + 
        labs(title="Luxury Car Models body-type", subtitle= "")+ xlab('count')+
        ylab('Body')+
        theme + theme(axis.text.x = element_text(angle = 0, hjust=0.5))

df <- cars_luxury %>% group_by(body) %>%
      summarise(Avg.SalePrice = mean(sellingprice))

p4<-ggplot(df, aes(y = reorder(body, Avg.SalePrice), x = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'black', fill='salmon')+
        labs(title="Luxury Cars Model: Body-type vs Sale price" )+ ylab('Body') +
        theme +
        theme(axis.text.x = element_text(angle = 0, hjust=0.5))
p1 + p2 + p4 + p3

```






### FORD popular models and highest selling price
```{r}
options(repr.plot.width = 16, repr.plot.height = 6)
df <- cars_ford %>% group_by(model) %>%
      summarise(Avg.SalePrice = mean(sellingprice))%>%
      arrange(desc(Avg.SalePrice))%>%
      top_n(15)

p1<-ggplot(df, aes(x = reorder(model, -Avg.SalePrice), y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'gold', alpha=0.6)+
        labs(title="Ford: Model vs Sale price" )+ xlab('Model')+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p1

df_ford <- cars_ford %>% group_by(model) %>% 
           summarise(freq = n()) %>% 
           arrange(desc(freq))%>%
           top_n(15)
    
p2<-ggplot(df_ford, aes(x = reorder(model,(-freq)), y = freq)) + 
        geom_bar(stat = 'identity', fill="#56B4E9", color='black') + 
        labs(title="Ford: Popular Models", subtitle= "")+
        xlab('Model')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p2
```






### FORD: Selling price and frequency based on Body
```{r}
df_ford <- cars_ford %>% group_by(body) %>% 
            summarise(freq = n())  
    
p3<-ggplot(df_ford, aes(x = reorder(body,(-freq)), y = freq)) + 
        geom_bar(stat = 'identity', color="#000666", fill='gold') + 
        labs(title="Ford: Popular Models", subtitle= "")+
        xlab('Body')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p3

df <- cars_ford %>% group_by(body) %>%
      summarise(Avg.SalePrice = mean(sellingprice))

p4<-ggplot(df, aes(x = reorder(body, -Avg.SalePrice), y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'gold', alpha=0.6)+
        labs(title="Ford: Body-type vs Sale price" )+ xlab('Body')+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p4

```





### Chevrolet: Popular models and Highest selling price model
```{r}
options(repr.plot.width = 16, repr.plot.height = 6)
df <- cars_chev %>% group_by(model) %>%
      summarise(Avg.SalePrice = mean(sellingprice))%>%
      arrange(desc(Avg.SalePrice))%>%
      top_n(15)

p1<-ggplot(df, aes(x = reorder(model, -Avg.SalePrice), y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'gold', alpha=0.6)+
        labs(title="Chev: Model vs Sale price" )+ xlab('Model')+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p1

df_chev <- cars_chev %>% group_by(model) %>% 
           summarise(freq = n()) %>% 
           arrange(desc(freq))%>%
           top_n(15)
    
p2<-ggplot(df_chev, aes(x = reorder(model,(-freq)), y = freq)) + 
        geom_bar(stat = 'identity', fill="#56B4E9", color='black') + 
        labs(title="Chev: Popular Models", subtitle= "")+
        xlab('Model')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p2

```


### Chevrolet:popular body type  and body type with highest selling price

```{r}
df_chev <- cars_chev %>% group_by(body) %>% 
            summarise(freq = n()) 
            
p3<-ggplot(df_chev, aes(x = reorder(body,(-freq)), y = freq)) + 
        geom_bar(stat = 'identity', color="#000666", fill='gold') + 
        labs(title="Chev: Popular Models", subtitle= "")+
        xlab('Body')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p3


df <- cars_chev %>% group_by(body) %>%
      summarise(Avg.SalePrice = mean(sellingprice))

p4<-ggplot(df, aes(x = reorder(body, -Avg.SalePrice), y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'gold', alpha=0.6)+
        labs(title="Chev: Body-type vs Sale price" )+ xlab('Body')+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))
p4
```






```{r}
options(repr.plot.width = 16, repr.plot.height = 6)
df <- cars %>% group_by(color) %>%
      summarise(Avg.SalePrice = mean(sellingprice))

ggplot(df, aes(x = reorder(color, -Avg.SalePrice), y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', color= 'gold', alpha=0.6)+
        labs(title="Color on Sale price" )+ xlab('color')+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))

cars %>% 
    group_by(color) %>% 
    summarise(count = n()) %>% 
    ggplot(aes(x = reorder(color,(-count)), y = count)) + 
        geom_bar(stat = 'identity', fill='salmon', color='black') + 
        labs(title="Popular car colors")+
        xlab('color')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))

```






### Interior:Sale price and popularity on Interior
```{r}
df <- cars %>% group_by(interior) %>%
      summarise(Avg.SalePrice = mean(sellingprice))

ggplot(df, aes(x = reorder(interior, -Avg.SalePrice), y = Avg.SalePrice)) + 
        geom_bar(stat= 'identity', fill='#00BFC4', color='black', alpha=0.6)+
        labs(title="Car's interior on Sale price" )+ xlab('interior')+
        theme +
        theme(axis.text.x = element_text(angle = 90, hjust=0.5))

cars %>% 
    group_by(interior) %>%
    summarise(count = n()) %>% 
    ggplot(aes(x = reorder(interior,(-count)), y = count)) + 
        geom_bar(stat = 'identity', fill='salmon', color='black') + 
        labs(title="Interior color popularity")+
        xlab('interior')+
        theme + theme(axis.text.x = element_text(angle = 90, hjust=0.5))
```






```{r}
options(repr.plot.width = 15, repr.plot.height = 4)
df <- cars %>% group_by(state) %>%
        summarise(Avg.SalePrice = mean(sellingprice))

ggplot(df, aes(x = reorder(state, -Avg.SalePrice), y = Avg.SalePrice,)) + 
        geom_bar(stat= 'identity', fill='#5F8790', color='gold', alpha=1.0)+
        labs(title="Avg. Sale price per state", xaxis='States' )+
        xlab('States')+
        theme +
        theme(axis.text.x = element_text(angle = 0, hjust=0.5))

cars %>% 
    group_by(state) %>% 
    summarise(freq = n()) %>% 
    ggplot(aes(x = reorder(state,(-freq)), y = freq)) + 
        geom_bar(stat = 'identity', fill='salmon', color='black') + 
        labs(title="Numbers of cars auctioned per state")+
        xlab('States')+ ylab('Count')+
        theme
```

